Estimation of the road traffic sound levels in urban areas based on non-negative matrix factorization techniques

### 2018-01-12 paper/gloaguenNmfTraffic.tex
done proof read

\subsubsection{Metrics}
The performance of the two estimators (filter and NMF) of the road traffic sound level are assessed through the calculation of one metric, the Mean Absolute Error ($MAE$). It expresses the quality of the long-term reconstruction of the signal and consists in the average over the $M$ sound scenes of the absolute difference between the exact and estimated traffic sound level in dB,

@@

In all, according Table \ref{tab:experimental_factors}, 24540 settings are performed between the different form of the dictionary $\mathbf{W}$ and the multiple experimental factors taken into account by NMF. For the filter estimator, between the $TIR$, the sub-classes and $f_c$, 180 settings are performed whereas for Sup NMF and Sem NMF, 3780 associations of factors are made. For TI NMF, 8400 combinations can be calculated.

@@

Table \ref{tab:results} summarized, according to the 2 main factors (\textit{method}, $\beta$), the $MAE$ error averaged on all sub-classes and all $TIR$ (750 sound mixtures in all). For the low-passa frequency filters and each NMF approaches, the best parameter combinations are detailed according to the $TIR$ in Table \ref{tab:results_TIR},  and are expand to the sub-classes in Figures \ref{fig:TIR_class_filter}, \ref{fig:TIR_class_sup}, \ref{fig:TIR_class_semi} and \ref{fig:TIR_class_TI}. \\

First, the errors produced by the filter are detailed. $f_c = 20 $ kHz is equivalent to consider all the sound mixtures without distinction between traffic and others sound sources. Consequently, in low $TIR$ (-12 dB and -6 dB), where traffic component is scarce, the error is more important than in high $TIR$ (6 dB and 12 dB) where the traffic component is predominant. $f_c = 500$ Hz is the cut-off frequency with the lower mean error obtained. It is then the first baseline to use to compare the performances of NMF.

In low $TIR$, for \textit{alert} and \textit{animals}, sub-classes composed of higher frequencies, this filter is efficient as it removes these frequency components. For the other sub-classes where low frequency contents are present (storm for \textit{climate}, voices in \textit{humans}, planes, tramway and train in \textit{transport} and ventilation noise in \textit{mechanics}), the filter considers all the energy located in the pass-band and then do not dissociate the traffic element from the other sound sources. The errors are then nearly all superior to 4 dB and are overestimate ; see Figure \ref{fig:alert_-12}. In opposite, in high $TIR$, the error is due to the energy removed from the traffic which has the consequence to underestimate the sound levels ; see Figure \ref{fig:alert_12}. The 500 Hz filter finds a balance between what it is put aside in low $TIR$ and what it is remained in high $TIR$. \\

@@

Compare to the filter errors, the choice of some NMF approaches make it possible to decrease the road traffic sound level estimation. The supervised approach is the only method that has an average error superior to the 500 Hz filter baseline. Finally, it is the most sophisticated methods (Sem NMF and TI NMF) that have better results. The lowest average error is obtained for TI NMF for $\beta$ = 1 and threshold $t$ = 0.42 with the dictionary factors $K$ = 200 and $w_t$ = 500 ms. On the other hand, the semi-supervised approach has a higher error but yet with a lower standard deviation.

@@

According to Table \ref{tab:results_TIR} and Figures \ref{fig:TIR_bar}, the behavior between the 3 versions of NMF differs. In the case of Sup NMF, it fails to improve the filtering performances despite good results in high $TIR$.  the error are too important for low $TIR$ and this for all the sub-classes. This method reveals to be to rigid as $\mathbf{W}$ is composed of fixed traffic spectrum. In the case of low $TIR$, in the aim to reduce the objective function, see eq. \ref{eq:min-D-WH}, traffic elements are used whatever the sound event in the  sound scene. The case of a scene of the \textit{alert} sub-class is presented in Figure \ref{fig:lp_alert_-12}. Here when the car horn sounds, some elements of $\mathbf{W}$ are activated and are then considered as traffic component generating a wrong estimation of the sound level. This behavior disappear when the traffic component become predominant to the \textit{interfering} class, see Figure \ref{fig:lp_alert_12}. Thus composed the dictionary of only fixed traffic spectrum is not a sufficient way to estimate correctly the traffic sound level, $\tilde{L}_{p,traffic}$.

Consequently, with the add of a mobile part in the dictionary, $\mathbf{W_r}$, the semi-supervised approach  allows a better consideration of the \textit{interfering} class in low $TIR$. It brings a significant decrease of the errors for low $TIR$ as it can be see in Figure \ref{fig:TIR_class_semi}. The $\mathbf{W_r}$ composition, in the case of an \textit{alert} sub-class, is displayed in Figure \ref{fig:Y_alert-12} and this shows that the interfering class is easily integrated. The first element is mainly composed of high frequency bands which correspond to the car horns of the scene. This composition impacted directly the traffic sound level estimation as it can be seen in Figure \ref{fig:lp_alert_-12} where the traffic basics are no longer activated when the car horn sounds.

@@

However, the degrees of freedom of Sem NMF are restrictive for high $TIR$ as the errors exceed 2 dB for all sub-classes and increase with $TIR$ = 6 dB and 12 dB. Indeed, in order to reduce $D(\mathbf{V} \vert \vert \mathbf{WH}$), without constraint, Sem NMF is free to include traffic components in $\mathbf{W_r}$. Consequently, this behavior decreases the quality of the reconstruction of the traffic component. In Figure \ref{fig:Y_alert_12} for $TIR=12$ dB, the high frequency components of car horn have disappeared for the benefit of low frequency content. Consequently, the traffic sound level estimation is then underestimate as in Figure \ref{fig:lp_alert_12}. \\

@@

Finally, TI NMF with a threshold fixed at $t = 0.54$ and $\beta$ = 2, offers the lowest average results (Table \ref{tab:results}). If, according to each $TIR$, this method does not propose the lowest error, it succeeds to be a compromise between the two others NMF. Unlike Sup NMF, where $\mathbf{W}$ is fixed, and Sem NMF, where only $\mathbf{W_r}$ is updated, TI NMF updates $\mathbf{W}$ entirely to be adjusted to the scene and to adapt to the different sound environments. The closest elements of the \textit{traffic} component defined in $\mathbf{W_0}$ are then extracted to deduce the traffic signal. In Figure \ref{fig:dist_-12_12}, the similarity $D_{\theta}(\mathbf{W_0}||\mathbf{W})$ is displayed for 3 sub-classes for $TIR$ = [-12, 12] dB.

@@

For low $TIR$, as the traffic sound class is not predominant, the final dictionary differs a lot from the $\mathbf{W_0}$. With the thresholding, only a reduce number of basis vectors are considered as traffic components. In comparison to supervised results, this approach reduce significantly the error for the \textit{human} and \textit{transport} sub-classes.  However, for \textit{climate} and \textit{mechanics}, the error stay important as these interfering classes have similar spectrum. On high $TIR$, as the traffic is the main sound source, the similarity of the initial dictionary and $\mathbf{W}$ is higher which allows retaining more elements as traffic components and then decrease the error (< 1 dB). The kept elements are then more suited to the scenes than a fixed dictionary. The error for these $TIR$ is then due to the thresholding. With a low threshold, it is possible to decrease the error (for $TIR$ = 12 dB, the average error on all the sub-classes is 0.22 ($\pm$ 0.08) dB with $t$ = 0.30).

@@

In this work, NMF, a source separation method, was used to estimate the road traffic sound level in urban sound mixtures. It is an adapted approach to these sound environments because it easily takes into account the overlap between the multiple sound sources present in the cities and is adapted to monophonic sensor networks. Different versions of NMF have been studied as a supervised and semi-supervised approach. On a large corpus of sound, the supervised approach proves to be too restrictive to be adapted to different sound environments whereas the semi-supervised approach has, on the contrary, too many degrees of freedom on the mobile dictionary $\mathbf{W_r}$, decreasing its performance especially when the traffic is predominant. The proposed new approach, named Threshold Initialized NMF, where $\mathbf{W}$ is initialized with road traffic spectrum, updated and where the traffic elements are extracted by hard thresholding, proposes the lowest average error. It allows a compromise between the other two approaches where a prior knowledge of the targeted sound source is given as in the Sup NMF, and offers an adaptability to the different sound mixtures by updating $ \mathbf{W}$ entirely.

This method has the advantage to keep a low calculation cost and then can be implemented in embedded sensors. Furthermore, this approach stay adapted for other sound sources, as whistling bird, depending on what spectrum are put in $\mathbf{W}$ and can be adapted for other application than the road traffic sound level estimation. Finally, the add of constraints as sparsness \cite{hoyer2004non} and smoothness \cite{virtanen_monaural_2007} can easily be implemented and taken into account.

In parallel, this experience allowed the design of a large sound database composed of multiples samples of urban sounds and sound mixtures that are made available\footnote{\url{https://sandbox.zenodo.org/record/176695#.Wk4ow3kiGos}} for research communities dedicated to the detection, separation and recognition tasks of urban sound sources.
The experience has been lead with the Matlab software. In order to be a reproducible experience, all the programs\footnote{\url{https://github.com/jean-remyGloaguen/article2017EstimationAmbiance}} used for this experience are available online.

>>>>

\subsubsection{Metrics}

The performance of the road traffic sound level estimator is assessed through the calculation of one reference metric, the Mean Absolute Error ($MAE$). It expresses the quality of the long-term reconstruction of the signal and consists in the average over the $M$ sound scenes of the absolute difference between the exact and estimated traffic sound level in dB,

@@

\ml{c'est les deux niveaux moyens sur toute la scène que l'on compare ?, quelle est la longueur des scenes ?}

Considering the experimental settings derived from the different modalities of each experimental factor described in Table \ref{tab:experimental_factors}, 24540 settings are performed between the different form of the dictionary $\mathbf{W}$ and the multiple experimental factors taken into account by NMF. For the filter estimator, between the $TIR$, the sub-classes and $f_c$, 180 settings are performed whereas for Sup NMF and Sem NMF, 3780 associations of factors are made. For TI NMF, 8400 combinations are computed.

@@

Table \ref{tab:results} summarized, according to the 2 main factors (\textit{method}, $\beta$), the $MAE$ error averaged on all sub-classes and all $TIR$ (750 sound mixtures in all). For the low-pass frequency filters and each NMF approaches, the best parameter combinations are detailed according to the $TIR$ in Table \ref{tab:results_TIR},  and are expanded to the sub-classes in Figures \ref{fig:TIR_class_filter}, \ref{fig:TIR_class_sup}, \ref{fig:TIR_class_semi} and \ref{fig:TIR_class_TI}. \\

First, the errors produced by the filter are detailed. $f_c = 20 $ kHz is equivalent to consider all the sound mixtures without distinction between traffic and others sound sources. Consequently, in low $TIR$ (-12 dB and -6 dB), where traffic component is scarce, the error is more important than in high $TIR$ (6 dB and 12 dB) where the traffic component is predominant. $f_c = 500$ Hz is the cut-off frequency with the lower mean error obtained. It is then considered as the baseline to compare the performances of NMF.

In low $TIR$, for \textit{alert} and \textit{animals}, sub-classes composed of higher frequencies, this filter is efficient as it removes these frequency components. For the other sub-classes where low frequency contents are present (storm for \textit{climate}, voices in \textit{humans}, planes, tramway and train in \textit{transport} and ventilation noise in \textit{mechanics}), the filter considers all the energy located in the pass-band and then do not dissociate the traffic element from the other sound sources. The errors are then nearly all superior to 4 dB and are overestimate, see Figure \ref{fig:alert_-12}. In opposite, in high $TIR$, the error is due to the energy removed from the traffic which has the consequence to underestimate the sound levels, see Figure \ref{fig:alert_12}. The 500 Hz filter finds a balance between what it is put aside in low $TIR$ and what it is remained in high $TIR$. \\

@@

Compare to the filter errors, the choice of some NMF approaches make it possible to decrease the road traffic sound level estimation. The supervised approach is the only method that has an average error superior to the 500 Hz filter baseline. The Sem NMF and TI NMF have better results. The lowest average error is obtained for TI NMF for $\beta$ = 1 and threshold $t$ = 0.42 with the dictionary factors $K$ = 200 and $w_t$ = 500 ms. On the other hand, the semi-supervised approach has a higher error but with a lower standard deviation.

@@

According to Table \ref{tab:results_TIR} and Figures \ref{fig:TIR_bar}, the behavior between the 3 versions of NMF differs. In the case of Sup NMF, it fails to improve the filtering performances despite good results in high $TIR$.  the error are too important for low $TIR$ and this for all the sub-classes. This method reveals to be too rigid as $\mathbf{W}$ is composed of fixed traffic spectrum. In the case of low $TIR$, in the aim to reduce the objective function, see eq. \ref{eq:min-D-WH}, traffic elements are used whatever the sound event in the  sound scene. The case of a scene of the \textit{alert} sub-class is presented in Figure \ref{fig:lp_alert_-12}. Here when the car horn sounds, some elements of $\mathbf{W}$ are activated and are then considered as traffic component generating a wrong estimation of the sound level. This behavior disappear when the traffic component become predominant to the \textit{interfering} class, see Figure \ref{fig:lp_alert_12}. Thus forcing the dictionary to be only traffic spectra is not a sufficient way to estimate correctly the traffic sound level, $\tilde{L}_{p,traffic}$.

Consequently, with the addition of a mobile part in the dictionary, $\mathbf{W_r}$, the semi-supervised approach  allows a better consideration of the \textit{interfering} class in low $TIR$. It brings a significant decrease of the errors for low $TIR$ as it can be see in Figure \ref{fig:TIR_class_semi}. The content of $\mathbf{W_r}$, in the case of an \textit{alert} sub-class is displayed in Figure \ref{fig:Y_alert-12}. The interfering class is easily integrated. The first element is mainly composed of high frequency bands which correspond to the car horns of the scene. This composition impacted directly the traffic sound level estimation as it can be seen in Figure \ref{fig:lp_alert_-12} where the elements of the dictionnary dedicated to the traffic are no longer activated when the car horn is active.

@@

However, the relatively high degree of freedom of Sem NMF are restrictive for high $TIR$ as the errors exceed 2 dB for all sub-classes and increase with $TIR$ = 6 dB and 12 dB. In order to reduce $D(\mathbf{V} \vert \vert \mathbf{WH}$), without constraint, Sem NMF is free to include traffic components in $\mathbf{W_r}$. Consequently, this behavior decreases the quality of the reconstruction of the traffic component. In Figure \ref{fig:Y_alert_12} for $TIR=12$ dB, the high frequency components of car horn have disappeared for the benefit of low frequency content. Consequently, the traffic sound level estimation is then underestimated as in \ml{rappler la condition experimentale}, see Figure \ref{fig:lp_alert_12}. \\

@@

Finally, TI NMF with a threshold $t = 0.54$ and $\beta$ = 2, offers the lowest average error (Table \ref{tab:results}). \ml{enlever ou reecrire: If, according to each $TIR$, this method does not propose the lowest error, it succeeds to be a compromise between the two others NMF.} Unlike Sup NMF, where $\mathbf{W}$ is fixed, and Sem NMF, where only $\mathbf{W_r}$ is updated, TI NMF updates $\mathbf{W}$ entirely to adjust prior knowledge to the scene under evaluation so as to adapt to the different sound environments. The closest elements of the \textit{traffic} component defined in $\mathbf{W_0}$ are then extracted to deduce the traffic signal. In Figure \ref{fig:dist_-12_12}, the similarity $D_{\theta}(\mathbf{W_0}||\mathbf{W})$ is displayed for 3 sub-classes and $TIR$ = [-12, 12] dB.

@@

For low $TIR$, as the traffic sound class is not predominant, the final dictionary stronglyb differs from the $\mathbf{W_0}$. With the thresholding, only a reduced number of basis vectors are considered as traffic components. In comparison to supervised results, this approach reduce significantly the error for the \textit{human} and \textit{transport} sub-classes.  However, for \textit{climate} and \textit{mechanics}, the error remain important as these interfering classes have similar spectral profiles when compared to traffic ones.

For high $TIR$, as the traffic is the main sound source, the similarity of the initial dictionary and $\mathbf{W}$ is higher which allows retaining more elements as traffic components and then decrease the error (< 1 dB). The kept elements are then more suited to the scenes than a fixed dictionary. The error for these $TIR$ is then due to the thresholding. With a low threshold, it is possible to decrease the error (for $TIR$ = 12 dB, the average error on all the sub-classes is 0.22 ($\pm$ 0.08) dB with $t$ = 0.30).

@@

In this work the non negative matric factorization framework  was used to estimate the road traffic sound level in urban sound mixtures. It is a well suited approach to these sound environments because it easily take into account the overlap between the multiple sound sources present in the cities and is adapted to monophonic sensor networks. Different versions of NMF have been studied as a supervised and semi-supervised approach. On a large corpus of sound, the supervised approach proves to be too restrictive to be adapted to different sound environments whereas the semi-supervised approach has, on the contrary, too many degrees of freedom on the mobile dictionary $\mathbf{W_r}$, decreasing its performance especially when the traffic is predominant. The proposed approach, named threshold initialized NMF, where $\mathbf{W}$ is initialized with road traffic spectrum, updated and where the dictionary elements that are similar to the traffic are extracted by hard thresholding, achieves the lowest average error. %It allows a compromise between the other two approaches where a prior knowledge of the targeted sound source is given as in the Sup NMF, and offers an adaptability to the different sound mixtures by updating $ \mathbf{W}$ entirely.

A major advantage of the proposed approach is that it is not designed for a specific source. Eventhough the experiments described in this paper focused on traffic sounds, changing the dictionarry to contain bird sounds will lead to an estimator of the presence of birds.

Extending the apporach to other sources is thus of intetrest for future research. Also, performance improvement could be achieved by the addition of constraints such as sparsness \cite{hoyer2004non} and smoothness \cite{virtanen_monaural_2007} of the low rank matrices.

The experimental protocol and the evaluated estimators has been implemented with the Matlab software. For reproducibility purposes, the code is available online\footnote{\url{https://github.com/jean-remyGloaguen/article2017EstimationAmbiance}}. The evaluation database composed of multiples samples of urban sounds and sound mixtures is also available\footnote{\url{https://sandbox.zenodo.org/record/176695#.Wk4ow3kiGos}} for the research community with interest in detection, separation and recognition tasks of urban sound sources.
